---
title: "Final Project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load required libraries

```{r}
library(gmodels)
library(C50)
```


## Import and clean data

```{r}
# Read in data from csv file
hotels <- read.csv("hotel_bookings.csv")

# Convert appropriate character variables to factors
hotels$hotel <- as.factor(hotels$hotel)
hotels$arrival_date_month <- as.factor(hotels$arrival_date_month)
hotels$market_segment <- as.factor(hotels$market_segment)
hotels$deposit_type <- as.factor(hotels$deposit_type)
hotels$customer_type <- as.factor(hotels$customer_type)
hotels$reserved_room_type <- as.factor(hotels$reserved_room_type)
hotels$distribution_channel <- as.factor(hotels$distribution_channel)
hotels$meal <- as.factor(hotels$meal)
hotels$assigned_room_type <- as.factor(hotels$assigned_room_type)

# Remove certain fields which provide no additional information
hotels$reservation_status <- NULL #duplicative of is_cancelled 
hotels$reservation_status_date <- NULL #may be of interest if we wanted to look at lead time of cancellation, but for our purposes we are removing
hotels$agent <- NULL #large factor of individual agent IDs - we should have whether or not an agent was involved (and type of agent) based on distribution_channel
hotels$company <- NULL #large factor of individual companies - we should have whether or not this is a corporate booking based on distribution_channel
hotels$country <- NULL #large factor of country codes - may be interesting if we think there is some sort of cultural impact on cancellations, but it is slowing down our model training to the point where it has become difficult to work with included
  

# Examine structure and summary of dataset
str(hotels)
summary(hotels)
```

## Create training and test sets for logit model and decision tree

```{r}
# Select 20,000 random rows for test data
set.seed(12345)
test_set <- sample(1:nrow(hotels), 20000)

# Create a train set and test set
hotels_train <- hotels[-test_set, ]
hotels_test <- hotels[test_set, ]
```


## First Logit Model
```{r}
# Starting with all potential independent variables
logit1 <- glm(is_canceled ~ ., data=hotels_train, family="binomial")
summary(logit1)

```
## Second Logit Model
```{r}
# Removing insignificant independent variables from logit1
logit2 <- glm(is_canceled ~ . -arrival_date_year -required_car_parking_spaces -days_in_waiting_list, data=hotels_train, family="binomial")
summary(logit2)
```

## Predicting test dataset with final logit model

```{r}
# Use logit2 to predict cancellations in hotels_test
logit_predict <- predict(logit2, hotels_test, type="response")
logit_predict <- ifelse(logit_predict > 0.5, 1, 0)

# View confusion matrix
CrossTable(hotels_test$is_canceled, logit_predict)
```
The logistic regression model is able to predict hotel room cancellations with an accuracy of **`r round(mean(logit_predict == hotels_test$is_canceled)*100)`**%.

## Decision Tree Model

```{r}
# Train decision tree using all available independent variables
tree1 <- C5.0(as.factor(is_canceled) ~ . , data=hotels_train)

# Note that we aren't displaying the decision tree itself, as the output is massive

# Use tree1 to predict cancellations in hotels_test
tree_predict <- predict(tree1, hotels_test)
#tree_predict <- ifelse(logit_predict > 0.5, 1, 0)

# View confusion matrix
CrossTable(hotels_test$is_canceled, tree_predict)

```
The decision tree model is able to predict hotel room cancellations with an accuracy of **`r round(mean(hotels_test$is_canceled == tree_predict)*100)`**%.

## Create all-numeric, normalized training and test datasets for ANN and KNN models
```{r}
# Using model.matrix to convert all the factors to dummy variables
hotels_numeric <- as.data.frame(model.matrix(~.-1,hotels))
str(hotels_numeric)
summary(hotels_numeric)

# Removing a dummy variable which gets created with the previous code which for some reason is all NAs
hotels_numeric$market_segmentUndefined <- NULL

# Create function to normalize data
normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}

# Apply normalization function to all-numeric dataset
hotels_numeric <- as.data.frame(lapply(hotels_numeric, normalize))

# Selects 20,000 random rows for test data
set.seed(12345)
test_set <- sample(1:nrow(hotels_numeric), 20000) 

# Create a train set and test set
hotels_numeric_train <- hotels_numeric[-test_set, ]
hotels_numeric_test <- hotels_numeric[test_set, ]
```
```{r}
hotels_numeric_trainlabels <- hotels_numeric_train[, "is_canceled"]
hotels_numeric_testlabels <- hotels_numeric_test[, "is_canceled"]

library(class)

hotelknn_predict <- knn(train = hotels_numeric_train, test = hotels_numeric_test, cl = hotels_numeric_trainlabels, k = 100)

CrossTable(x = hotels_numeric_testlabels, y = hotelknn_predict, prop.chisq = FALSE)
```


